\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-10-18 Documents/M1/Astronum_2018/manuscript.tex
Backup the modification made for abstract, introduction and two-moment method section.

\begin{abstract}
We develop implicit-explicit (IMEX) schemes for neutrino transport through a background material in the \texttt{thornado} code in the context of a two-moment mode that evolves the angular moments of a neutrino phase-space distribution function.
Considering the upper and lower bounds that are introduced by Pauli's exclusion principle on the moments, an algebraic moment closure based on Fermi-Dirac statistics and a convex-invariant time integrator are demanded.
Several algebraic closures are compared with these bounds in mind and the Cernohorsky and Bludman closure\cite{cernohorskyBludman_1994}, which preserves the bounds, is chosen for our IMEX schemes.
For the convex-invariant time integrator, two IMEX scheme named PD-ARS have been proposed following the work in \cite{chu_etal_2018}.
PD-ARS denotes a convex-invariant IMEX Runge-Kutta scheme that has first-order accuracy in general, high-order accuracy in the streaming limit, and works well in the diffusion limit.
Our two PD-ARS schemes are second- and third-order, explicit, strong-stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} as their explicit part, respectively, and therefore are second- and third-order accurate in the streaming limit, respectively.
The accuracy and convex-invariance of our PD-ARS schemes are demonstrated in the numerical tests with a first-order discontinuous Galerkin(DG) method for spatial discretization and a simple Lax-Friedrichs flux.

>>>>

\begin{abstract}
We develop implicit-explicit (IMEX) schemes for neutrino transport through a background material in the \texttt{thornado} code in the context of a two-moment model that evolves the angular moments of a neutrino phase-space distribution function.
Considering the upper and lower bounds that are introduced by Pauli's exclusion principle on the moments, an algebraic moment closure based on Fermi-Dirac statistics and a convex-invariant time integrator are demanded.
A finite-volume/first-order discontinuous Galerkin(DG) method is used to interporate how an algebraic moment closure based on Fermi-Dirac statistics satisfies the bounds.
Several algebraic closures are compared with these bounds in mind and the Cernohorsky and Bludman closure\cite{cernohorskyBludman_1994}, which satisfies the bounds, is chosen for our IMEX schemes.
For the convex-invariant time integrator, two IMEX scheme named PD-ARS have been proposed following the work in \cite{chu_etal_2018}.
PD-ARS denotes a convex-invariant IMEX Runge-Kutta schemes that has first-order accuracy in general, high-order accuracy in the streaming limit, and works well in the diffusion limit.
Our two PD-ARS schemes are second- and third-order, explicit, strong-stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} as their explicit part, respectively, and therefore are second- and third-order accurate in the streaming limit, respectively.
The accuracy and convex-invariance of our PD-ARS schemes are demonstrated in the numerical tests with a third-order DG method for spatial discretization and a simple Lax-Friedrichs flux.
We show preliminary result employing tabulated neutrino opacities from Bruenn(1985)\citep{Bruenn_1985}

### 2018-10-18 Documents/M1/Astronum_2018/sections/TimeIntegration.tex
Backup the modification made for abstract, introduction and two-moment method section.

It can be proven that the conditions we need for a simple convex-invariant IMEX scheme are: \cite{chu_etal_2018}
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item Second-order accuracy in the streaming limit:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    And $c_{ij}$ and $\hat{c}_{ij}$ be given by
    \begin{equation}
     \begin{aligned}
      c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
      c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj} \\
      \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
      \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}.  
     \end{aligned}
     \label{eq:positivityCoefficientsARS}
    \end{equation}
    \begin{equation}
      b_{ii} = \f{1}{a_{ii}}, \quad
      b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
      \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
    \end{equation}
    Note that $c_{i1}=\tilde{c}_{i1}=0$ in Eq.~\eqref{eq:positivityCoefficientsARS} so that $\sum_{j=0}^{i-1}c_{ij}=1$.
    \item Well-behaved in the diffusion limit:
    \begin{equation}
      \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s,
      \label{eq:diffusionCondition}
    \end{equation}
    where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix.
    \item Less than four stages ($s\le3$).
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. (GSA)
\end{enumerate}  
We call the IMEX scheme satisfying the above conditions {PD-ARS}. (Definition 3 in \cite{chu_etal_2018}.)
We give two optimized PD-ARS schemes here: PD-ARS with SSPRK2 and  PD-ARS with SSPRK3. 
They have second- and third-order accuracy in the streaming limit, respectively.
\subsubsection{PD-ARS with SSPRK2}
Here we give a family of PD-ARS with SSPRK2 in the standard double Butcher tableau form: explicit tableau on the left ($\tilde{A}$), implicit tableau on the right ($A$):
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2-\epsilon & 1/2+\epsilon \\ \hline
  	  & 0 & 1/2-\epsilon & 1/2+\epsilon
  \end{array}
\end{equation}
and $\epsilon \in [0, 1/2)$.
For these schemes, $c_{\mbox{\tiny Sch}}= 1 - 2\epsilon$.
The optimized scheme ($c_{\mbox{\tiny Sch}}= 1$) is given by $\epsilon = 0$.
We chose $\epsilon = 0.1$ for the numerical test in Section~\ref{se:NumericalTests}.

>>>>

It can be proven that the conditions we need for a simple convex-invariant IMEX scheme are: \cite{chu_etal_2018}
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item Second-order accuracy in the streaming limit:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{i1} = 0, \quad \text{for} \quad i=1,\ldots,s, \nonumber \\
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    And $c_{ij}$ and $\hat{c}_{ij}$ be given by
    \begin{equation}
     \begin{aligned}
      c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
      c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj} \\
      \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
      \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}.  
     \end{aligned}
     \label{eq:positivityCoefficientsARS}
    \end{equation}
    \begin{equation}
      b_{ii} = \f{1}{a_{ii}}, \quad
      b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
      \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
    \end{equation}
    Note that $c_{i1}=\tilde{c}_{i1}=0$ in Eq.~\eqref{eq:positivityCoefficientsARS} so that $\sum_{j=0}^{i-1}c_{ij}=1$.
    \item Well-behaved in the diffusion limit:
    \begin{equation}
      \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s,
      \label{eq:diffusionCondition}
    \end{equation}
    where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix.
    \item Less than four stages ($s\le3$).
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. (GSA)
\end{enumerate}  
We call the IMEX scheme satisfying the above conditions {PD-ARS}. (Definition 3 in \cite{chu_etal_2018}.)
We give two optimized PD-ARS schemes here: PD-ARS with SSPRK2 and  PD-ARS with SSPRK3. 
They have second- and third-order accuracy in the streaming limit, respectively.
\subsubsection{PD-ARS with SSPRK2}
Here we give a family of PD-ARS with SSPRK2 in the standard double Butcher tableau form: explicit tableau on the left ($\tilde{A}$), implicit tableau on the right ($A$):
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2-\epsilon & 1/2+\epsilon \\ \hline
  	  & 0 & 1/2-\epsilon & 1/2+\epsilon
  \end{array}
\end{equation}
and $\epsilon \in [0, 1/2)$.
For these schemes, $c_{\mbox{\tiny Sch}}= 1 - 2\epsilon$.
The optimized scheme ($c_{\mbox{\tiny Sch}}= 1$) is given by $\epsilon = 0$.
We chose $\epsilon = 0.1$ for the numerical test in Section~\ref{se:NumericalTests}.

### 2018-10-18 Documents/M1/Astronum_2018/sections/TwoMoments.tex
Backup the modification made for abstract, introduction and two-moment method section.

\subsection{Transport Equations}
Consider neutrino transport through static dense matter with only emission, absorption and isotropic scattering.
The Boltzmann equation after scaling to dimensionless units ($k_{B} = \hbar = c = 1$) can be written as
\begin{equation}
  \pd{f}{t}+\vect{\ell}\cdot\nabla f
  =\f{1}{\tau}\,\cC(f).
  \label{eq:boltzmann}
\end{equation}
The distribution function $f = f(\omega,\varepsilon,\vect{x},t)$ gives the number of neutrinos propagating in the direction $\omega\in\bbS^{2}$, with neutrino energy $\varepsilon\in\bbR^{+}$, at position $\vect{x}\in\bbR^{3}$ and time $t\in\bbR^{+}$.  
$\vect{\ell} = \vect{\ell}(\omega)\in\bbR^{3}$ is the unit vector parallel to the neutrino three-momentum direction: $\vect{p}=\varepsilon\,\vect{\ell}$.
On the right-hand side, $\tau$ is the interaction strength parameter: $\tau\ll1$ for opaque regions where neutrinos have frequent interactions with the background; $\tau\gg1$ for transparent regions where neutrinos rarely interact with the background and stream freely.
$\cC(f)$ is the collision term, which models emission, absorption, and isotropic and elastic scattering: 
\begin{equation}
  \cC(f)=\xi\,\big(\,f_{0}-f\,\big)
  +(1-\xi)\,\big(\,\f{1}{4\pi}\int_{\bbS^{2}}f\,d\omega-f\,\big),
  \label{eq:collisionTerm}
\end{equation}
with $\xi = \sigma_{\Ab} / (\sigma_{\Ab}  + \sigma_{\Scatt} )$: $\xi = 1$ when the scattering opacity $\sigma_{\Scatt} = 0$ corresponding to pure emission and absorption; $\xi = 0$ when the absorption opacity $\sigma_{\Ab} = 0$ corresponding to pure scattering. 
$f_{0}$ is the neutrino equilibrium distribution function, which has the following form:
\begin{equation}
  f_{0}(\vect{z})=\f{1}{e^{(\varepsilon-\mu(\vect{x}))/T(\vect{x})}+1}.
  \label{eq:fermiDirac}
\end{equation}
$T(\vect{x})$ is the temperature in energy units and $\mu(\vect{x})$ is the neutrino chemical potential.
Both of them depend on the properties of the background as a function of $\vect{x}$.

\subsection{Two-Moment Model}
An approximate solution of the Boltzmann equation, Eq.~\eqref{eq:boltzmann}, can be found by employing a two-moment model.
Define the angular moments of the distribution function as follows
\begin{equation}
  \big\{\,\cJ,\vect{\cH},\vect{\cK}\,\big\}(\vect{z},t)
  =\f{1}{4\pi}\int_{\bbS^{2}}f(\omega,\vect{z},t)\,\{\,1,\vect{\ell},\vect{\ell}\otimes\vect{\ell}\,\}\,d\omega,
  \label{eq:angularMoments}
\end{equation}
where $\vect{z}:=\{\varepsilon,\vect{x}\}$.
The zeroth moment, $\cJ$, is referred to as the particle density.
The first moment, $\bcH$, is the particle flux.
And the second moment, $\bcK$, is the stress tensor.
We can integrate Eq.~\eqref{eq:boltzmann} to their equations of the zeroth and the first moments
\begin{equation}
  \pd{\vect{\cM}}{t}+\nabla\cdot\vect{\cF}=\f{1}{\tau}\,\vect{\cC}(\vect{\cM}),
  \label{eq:momentEquations}
\end{equation}
with $\vect{\cM}=(\cJ,\vect{\cH})^{T}$, $\vect{\cF}=(\vect{\cH},\vect{\cK})^{T}$, and
\begin{equation}
  \vect{\cC}(\vect{\cM})=\vect{\eta}-\vect{\cD}\,\vect{\cM},
  \label{eq:collisionTermMoments}
\end{equation}
where $\vect{\eta}=(\xi\,f_{0},\vect{0})^{T}$, $\vect{\cD}=\mbox{diag}(\xi,\vect{I})$, and
$\vect{I}$ is the identity matrix.
Therefore, solving the Boltzmann equation, Eq.~\eqref{eq:boltzmann}, for the neutrino distribution function, $f(\omega,\vect{z},t)$, is approximated by solving the two-moment equations for the neutrino number density, $\cJ(\vect{z},t)$, and flux, $\bcH(\vect{z},t)$.

\subsection{Algebraic Closures }
As is shown in the previous section, the $i$-th order moment equation requires information of the $(i+1)$-th order moment and the two-moment system is open. 
To close the two-moment system, an algebraic closure is constructed.
For a two-moment method, algebraic closures give an approximate $\bcK$ based on the lower moments as follows:
\begin{equation}
\bcK = \vect{k} \cJ,
\end{equation}
where $\vect{k}$ is the Eddington tensor.
For example, assuming the neutrino transport is symmetric about a preferred direction $\widehat{\vect{h}}=\vect{\cH}/|\vect{\cH}|$, Levermore\cite{levermore_1984} proposed
\begin{equation}
  \vect{k}=\f{1}{2}\big[\,\big(1-\chi\big)\,\vect{I}+\big(3\,\chi-1\big)\,\widehat{\vect{h}}\otimes\widehat{\vect{h}}\,\big],
  \label{eq:eddingtonTensor}
\end{equation}
where $\chi=\chi(\cJ,|\vect{\cH}|)$ is the Eddington factor.

\subsection{Constraints on the Moments}
One fundamental fact about neutrinos is that they are fermions, which obey follow the Pauli exclusion principle, $f \in [0,1]$.
As a result, the angular moments of $f$ are constrained as follows: 
\begin{align}
\cJ \in[0,1], \quad &(1-\cJ)\cJ-|\vect{\cH}|  \geq 0, \label{eq:MomentsBounds}\\
  \chi_{\mbox{\tiny min}}
  =\max\big(1-\f{2}{3\cJ},h^{2}\big)
  \leq & \chi \leq \min\big(1,\f{1}{3\cJ}-\f{\cJ}{1-\cJ}h^{2}\big)=\chi_{\mbox{\tiny max}},
  \label{eq:eddingtonFactorBounds}
\end{align}
where $h = |\bcH|/\cJ$ is the flux factor.\cite{levermore_1984,lareckiBanach_2011,kershaw_1976,shohatTamarkin_1943}

The inequalities in Eq.~\eqref{eq:MomentsBounds} define the realizable $\bcM$, which define a convex set in $\cJ -\bcH$ space.
As we will see later in Section \ref{se:SpacialDiscretization}, this convexity makes a constraint-preserving spatial discretization possible.

The inequalities in Eq.~\eqref{eq:eddingtonFactorBounds} deserve more attention than has been given.
They have equal importance \rc{relative to} the inequalities in Eq.~\eqref{eq:MomentsBounds} in preserving the validity of the simulation result.
However, as Fig.~\eqref{fig:EddingtonFactorsWithDifferentClosure} shows, not all algebraic closures satisfy the Eddington factor bounds, Eq.~\eqref{eq:eddingtonFactorBounds}: Kershaw\cite{kershaw_1976}, Wilson\cite{wilson_1975,leblancWilson_1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 2\cite{janka_1992} may work fine in the neutrino low occupancy limit, but not for high occupancies; Janka 1\cite{janka_1991} can be in error in both cases.
The only algebraic closure that remains within the Eddington factor bounds among those plotted is Cernohorsky and Bludman's \cite{cernohorskyBludman_1994}.

>>>>

\subsection{Transport Equations}
Consider neutrino transport through static dense matter with only emission, absorption and isotropic iso-energetic scattering.
The Boltzmann equation after scaling to dimensionless units ($k_{B} = c = 1$) can be written as
\begin{equation}
  \pd{f}{t}+\vect{\ell}\cdot\nabla f
  =\f{1}{\tau}\,\cC(f).
  \label{eq:boltzmann}
\end{equation}
The distribution function $f = f(\omega,\varepsilon,\vect{x},t)$ gives the number of neutrinos propagating in the direction $\omega\in\bbS^{2}$, with energy $\varepsilon\in\bbR^{+}$, at position $\vect{x}\in\bbR^{3}$ and time $t\in\bbR^{+}$.  
$\vect{\ell} = \vect{\ell}(\omega)\in\bbR^{3}$ is the unit vector parallel to the neutrino three-momentum direction: $\vect{p}=\varepsilon\,\vect{\ell}$.
On the right-hand side, $\tau$ is collision time scale: $\tau\ll1$ for opaque regions where neutrinos have frequent interactions with the background; $\tau\gg1$ for transparent regions where neutrinos rarely interact with the background and stream freely.
$\cC(f)$ is the collision term, which models emission, absorption, and isotropic and elastic scattering: 
\begin{equation}
  \cC(f)=\xi\,\big(\,f_{0}-f\,\big)
  +(1-\xi)\,\big(\,\f{1}{4\pi}\int_{\bbS^{2}}f\,d\omega-f\,\big),
  \label{eq:collisionTerm}
\end{equation}
with $\xi = \sigma_{\Ab} / (\sigma_{\Ab}  + \sigma_{\Scatt} )$: $\xi = 1$ when the scattering opacity $\sigma_{\Scatt} = 0$ corresponding to pure emission and absorption; $\xi = 0$ when the absorption opacity $\sigma_{\Ab} = 0$ corresponding to pure scattering. 
$f_{0}$ is the neutrino equilibrium distribution function, which has the following form:
\begin{equation}
  f_{0}(\vect{z})=\f{1}{e^{(\varepsilon-\mu(\vect{x}))/T(\vect{x})}+1}.
  \label{eq:fermiDirac}
\end{equation}
$T(\vect{x})$ is the temperature in energy units and $\mu(\vect{x})$ is the neutrino chemical potential.
Both of them depend on the properties of the background as a function of $\vect{x}$.

\subsection{Two-Moment Model}
An approximate solution of the Boltzmann equation, Eq.~\eqref{eq:boltzmann}, can be found by employing a two-moment model.
Define the angular moments of the distribution function as follows
\begin{equation}
  \big\{\,\cJ,\vect{\cH},\vect{\cK}\,\big\}(\vect{z},t)
  =\f{1}{4\pi}\int_{\bbS^{2}}f(\omega,\vect{z},t)\,\{\,1,\vect{\ell},\vect{\ell}\otimes\vect{\ell}\,\}\,d\omega,
  \label{eq:angularMoments}
\end{equation}
where $\vect{z}:=\{\varepsilon,\vect{x}\}$.
The zeroth moment, $\cJ$, is referred to as the particle density.
The first moment, $\bcH$, is the particle flux, and the second moment, $\bcK$, is the stress tensor.
We can integrate Eq.~\eqref{eq:boltzmann} to obtain equations for the zeroth and the first moments
\begin{equation}
  \pd{\vect{\cM}}{t}+\nabla\cdot\vect{\cF}=\f{1}{\tau}\,\vect{\cC}(\vect{\cM}),
  \label{eq:momentEquations}
\end{equation}
with $\vect{\cM}=(\cJ,\vect{\cH})^{T}$, $\vect{\cF}=(\vect{\cH},\vect{\cK})^{T}$, and
\begin{equation}
  \vect{\cC}(\vect{\cM})=\vect{\eta}-\vect{\cD}\,\vect{\cM},
  \label{eq:collisionTermMoments}
\end{equation}
where $\vect{\eta}=(\xi\,f_{0},\vect{0})^{T}$, $\vect{\cD}=\mbox{diag}(\xi,\vect{I})$, and
$\vect{I}$ is the identity matrix.
Therefore, solving the Boltzmann equation, Eq.~\eqref{eq:boltzmann}, for the neutrino distribution function, $f(\omega,\vect{z},t)$, is approximated by solving the two-moment equations for the neutrino number density, $\cJ(\vect{z},t)$, and flux, $\bcH(\vect{z},t)$.

\subsection{Algebraic Closures }
As is shown in the previous section, the moment equation for $\bcH$ involves $\bcK$ and the two-moment system is open. 
To close the two-moment system, an algebraic closure is constructed.
For a two-moment method, algebraic closures give an approximate $\bcK$ based on the lower moments as follows:
\begin{equation}
\bcK = \vect{k} \cJ,
\end{equation}
where $\vect{k}$ is the Eddington tensor.
For example, assuming the neutrino distribution is symmetric about a preferred direction $\widehat{\vect{h}}=\vect{\cH}/|\vect{\cH}|$, Levermore\cite{levermore_1984} proposed
\begin{equation}
  \vect{k}=\f{1}{2}\big[\,\big(1-\chi\big)\,\vect{I}+\big(3\,\chi-1\big)\,\widehat{\vect{h}}\otimes\widehat{\vect{h}}\,\big],
  \label{eq:eddingtonTensor}
\end{equation}
where $\chi=\chi(\cJ,|\vect{\cH}|)$ is the Eddington factor.

\subsection{Constraints on the Moments}
Neutrinos are fermions, and obey the Pauli exclusion principle, i.e. $f \in [0,1]$.
As a result, the angular moments of $f$ and the Eddington factor $\chi$ are constrained as follows\cite{levermore_1984,lareckiBanach_2011,kershaw_1976,shohatTamarkin_1943}: 
\begin{align}
\cJ \in[0,1], \quad &(1-\cJ)\cJ-|\vect{\cH}| \geq 0, \label{eq:MomentsBounds}\\
  \chi_{\mbox{\tiny min}}
  =\max\big(1-\f{2}{3\cJ},h^{2}\big)
  \leq & \chi \leq \min\big(1,\f{1}{3\cJ}-\f{\cJ}{1-\cJ}h^{2}\big)=\chi_{\mbox{\tiny max}},
  \label{eq:eddingtonFactorBounds}
\end{align}
where $h = |\bcH|/\cJ$ is the flux factor.

The constraints in Eq.~\eqref{eq:MomentsBounds} define realizable $\bcM$, which define a convex set in the $\cJ$-$\bcH$ space.
As we will see later in Section \ref{se:SpacialDiscretization}, this convexity makes a constraint-preserving spatial discretization possible.

The inequalities in Eq.~\eqref{eq:eddingtonFactorBounds} deserve more attention than has been given.
They have equal importance relative to the inequalities in Eq.~\eqref{eq:MomentsBounds} in maintain consistency with respect to Fermi-Dirac statistics.

The Eddington factor given by the algebraic closures discussed in \cite{murchikova_etal_2017} are examined in this paper as examples.
We consider various occupancies and plot at $\cJ = 0.1$ and $\cJ = 0.9$ as examples of low occupancy and high occupancy, respectively, in Fig.~\eqref{fig:EddingtonFactorsWithDifferentClosure}.  
As the figure shows, not all algebraic closures satisfy the Eddington factor bounds, Eq.~\eqref{eq:eddingtonFactorBounds}: Kershaw\cite{kershaw_1976}, Wilson\cite{wilson_1975,leblancWilson_1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 2\cite{janka_1992} may work fine in the neutrino low occupancy limit, but not for high occupancies; Janka 1\cite{janka_1991} can be in error in both cases.
The only algebraic closure that remains within the Eddington factor bounds among those plotted is Cernohorsky and Bludman's \cite{cernohorskyBludman_1994}.
